---
title: "Analyzing Phrases with tidytext"
output:
  word_document: default
  html_notebook: default
  html_document: default
---

As always, we set the working directory and then load the packages we want.
```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(tidytext)
library(okcupiddata)
library(scales) # scale functions for visualization
library(igraph)
library(ggraph)
```

We continue the analysis of the OKCupid essays, and just to add some further interests we compare responses of smokers and non-smokers. 

For clarity, let's break the process down into steps. First, subset the okcupid `profiles` to include just the essay responses and the user's response about smoking.  The `data_frame` function creates a tibble. In creating the tibble, we rename `essay0` as `text` -- not necessary, but follows the example in Silge and Robinson.

Then we create a new binary variable, `smoker` that identifies non-smokers as "no" and everyone else as "yes", dropping the original `smokes` variable (notice the command `select(-smokes)`, which selects all columns *except for* `smokes`. 

```{r}
tidy_okcupid <- select_(profiles,"essay0","smokes") 

tidy_okcupid <- data_frame(smokes=profiles$smokes,text=profiles$essay0)
tidy_okcupid <- tidy_okcupid %>%
     mutate(smoker = ifelse(smokes=="no","no","yes")) %>% 
     select(-smokes)
```
The next few lines of code *remove* all `NA` rows, because the `unnest_tokens` function requires complete cases. The last line in this chunk unnests 2-word phrases from the . You may also want to experiment with longer phrases.

```{r}
tidy_smoker <- tidy_okcupid %>%
     select(smoker, text) %>%
     na.omit()

smoker_bigrams <- unnest_tokens(tidy_smoker, bigram, text, token="ngrams", n=2)

smoker_bigrams
```
At this point, the tibble `smoker_bigrams` consists of more than 1,226,000 pairs of adjacent words. The next code chunk counts and sorts the pairs to find the most common ones. After that, it `separate`s the pairs into `word1` and `word2` to search for and remove stopwords, and then `unite`s the pairs that remain, finally recounting and sorting the shorter list of pairs.

Consult Silge & Robinson to learn how to customize the list of stopwords.

```{r}
smoker_bigrams %>%
     count(bigram, sort = TRUE)
     
bigrams_separated <- smoker_bigrams %>%
     separate(bigram, c("word1", "word2"), sep = " ")
     
bigrams_filtered <- bigrams_separated %>%
     filter(!word1 %in% stop_words$word) %>%
     filter(!word2 %in% stop_words$word)
     
# new bigram counts:
bigram_counts <- bigrams_filtered %>% 
     count(word1, word2, smoker, sort = TRUE)
     

bigram_counts
     
bigrams_united <- bigrams_filtered %>%
     unite(bigram, word1, word2, sep = " ")
     
bigrams_united
```
As noted in class and in the readings, a simple frequency count can be misleading. The next code chunk weights term frequency by the inverse frequency within the documents (smokers vs non-smokers.
```{r}
     bigram_tf_idf <- bigrams_united %>%
          count(smoker, bigram) %>%
          bind_tf_idf(bigram, smoker, n) %>%
          arrange(desc(tf_idf))
     
     bigram_tf_idf

     bigrams_separated %>%
          filter(word1 == "not") %>%
          count(word1, word2, sort = TRUE)
```
```{r}
plot_pairs <- bigram_tf_idf %>%
  arrange(desc(tf_idf)) %>%
  mutate(pair = factor(bigram, levels = rev(unique(bigram))))

plot_pairs %>% 
  group_by(smoker) %>% 
  top_n(15) %>% 
  ungroup %>%
  ggplot(aes(pair, tf_idf, fill = smoker)) +
  geom_col(show.legend = FALSE) +
  labs(x = NULL, y = "tf-idf") +
  facet_wrap(~smoker, ncol = 2, scales = "free") +
  coord_flip()
```


# Sentiment Analysis

DataCamp introduced you to Sentiment Analysis. This next chunk follows the Silge & Robinson example. We specifically look for word pairs beginning that include negation (like "not"), since that can be a common style.

```{r}
     AFINN <- get_sentiments("afinn")
     
     not_words <- bigrams_separated %>%
          filter(word1 == "not") %>%
          inner_join(AFINN, by = c(word2 = "word")) %>%
          count(word2, score, sort = TRUE) %>%
          ungroup()
     
     not_words
     
     not_words %>%
          mutate(contribution = n * score) %>%
          arrange(desc(abs(contribution))) %>%
          head(20) %>%
          mutate(word2 = reorder(word2, contribution)) %>%
          ggplot(aes(word2, n * score, fill = n * score > 0)) +
          geom_col(show.legend = FALSE) +
          xlab("Words preceded by \"not\"") +
          ylab("Sentiment score * number of occurrences") +
          coord_flip()
     negation_words <- c("not", "no", "never", "without")
     
     negated_words <- bigrams_separated %>%
          filter(word1 %in% negation_words) %>%
          inner_join(AFINN, by = c(word2 = "word")) %>%
          count(word1, word2, score, sort = TRUE) %>%
          ungroup()
```
At this point we begin to make network graphs, first showing the most common word combinations for *non-smokers. You could then repeat and modify the chunk for smokers*.

```{r}
     
     # original counts
     bigram_counts
     
# filter for only relatively common combinations
# might want to experiment with values other than 20 in filter line
     bigram_graph <- bigram_counts %>%
          group_by(smoker) %>%
          filter(n > 20) %>%
          filter(smoker=="yes") %>%
          graph_from_data_frame()
     
     bigram_graph

     set.seed(2018)
     
     ggraph(bigram_graph, layout = "fr") +
          geom_edge_link() +
          geom_node_point() +
          geom_node_text(aes(label = name),  vjust = 1, hjust = 1) 
```
Finally, as in Chapter 4 of Silge and Robinson (which was not assigned), here's an alternative and more attractive network graph. This will use the same subset of participants as the one created above.

```{r}
     set.seed(2018)
     
     a <- grid::arrow(type = "closed", length = unit(.05, "inches"))
     
     ggraph(bigram_graph, layout = "fr") +
          geom_edge_link(aes(edge_alpha = n), show.legend = FALSE,
                         arrow = a, end_cap = circle(.05, 'inches')) +
          geom_node_point(color = "lightblue", size = 5) +
          geom_node_text(aes(label = name), vjust = 1, hjust = 1) +
          theme_void()
     


```

